---
title: "Hands-on with synapser and datasets"
output: html_document
date: "2024-06-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(synapser)
synLogin() 

# Note: can also do synLogin(authToken="your-token") but not recommended in case 
# this gets shared with others later or stored in version control 

tutorial_project <- "syn60667554"
```

## Using an analysis-ready dataset

The example dataset is "Copy Number Variation from the Columbia University NF1 Glioma Project" (syn51915976).

### Get metadata in the dataset

Use SQL-like query to get metadata / basic clinical data as table.

```{r query}

dataset_id <- "syn51915976"
dataset_version <- "4"
dataset_id_with_version <- "syn51915976.4"
query <- glue::glue("SELECT id as fileId,name as fileName,individualID,specimenID,diagnosis,nf1Genotype,nf2Genotype,sex,age,grade,histology FROM {dataset_id_with_version}")
query

```

```{r the-table}
results <- synTableQuery(query)
dt <- as.data.frame(results)
dt
```

### Batch download files in a dataset

First way to download the dataset -- use your download list if you've built one in Synapse UI.

```{r download-dataset-as-download-list}

download1 <- synGetDownloadList(downloadLocation = "./download1")
```

Second way to download a dataset -- get ids of the items and download via a loop.

```{r files-in-dataset}

reused_dataset <- synGet(dataset_id) 
file_list <- reused_dataset$properties$datasetItems
head(file_list)

```

```{r download-files}

for (i in file_list) {
  synGet(entity = i$entityId, 
         version = i$versionNumber,
         downloadFile = TRUE, 
         downloadLocation = "./download2")
}
```

## Creating a dataset

When and why:
- Analysis-ready dataset doesn't already exist for some collection of files.
- Create a dataset in your own project that will keep track with each analysis version.
Often an analysis will be done a number of times with:
  - Removal of some data.
  - Different versions of some data.
  - Add/combine more data, maybe from another project.

```{r first-dataset}

# Dataset constructor
dataset1 <- synapser::Dataset(
  name = "Dataset 1 - Data From One Project", 
  parent = tutorial_project,
  folders = list("syn18407530"), 
  addAnnotationColumns = TRUE)  # will surface all available metadata, TRUE is a good default in most cases

# Store new dataset in Synapse
dataset1 <- synStore(dataset1)

```


## Batch Download Files From Your Dataset

Once you have create and stored the dataset, can use item ids for download and **batch download exactly as shown above**.

Get the ids of files (i.e. "dataset items") to download.
```{r files-in-dataset}

file_list_2 <- dataset1$properties$datasetItems
head(file_list_2)

```

